---
title: "Titanic Dataset - Exploration with R"
output: html_notebook
---

Get the required packages:
```{r}
require(ggplot2)
```

Load the data into R:
```{r}
train <- read.csv("train.csv", stringsAsFactors = F) #For cleaning purposes let's not convert strings to factors.
test <- read.csv("test.csv", stringsAsFactors = F)
```

Let's take a look at what is in the train and test data sets.
```{r}
str(train)
```
```{r}
train$Pclass <-  as.factor(train$Pclass)
```

Let's see how many missing values there are in each of the columns.
```{r}
for (n in names(train)){
  cat(sprintf("%s missing %i values.\n", n, sum(is.na(train[n]))))
}
```
It appears that the "Age" column has a significant number of missing values.
Luckily however there are not missing values in the other columns.

Before we remove or impute the missing values, let's just make sure there is not a correlation between the missing values and the predictions.
```{r}
#Let's get a baseline for what ratio survived.
sum(train$Survived)/length(train$Survived)
#Now do the same for the "NA" Ages.
train_na <- subset(train, is.na(train$Age))
sum(train_na$Survived)/length(train_na$Survived)
```
It doesn't appear that there is a huge difference between the number of missing ages and the survival rate.

Since this isn't a huge dataset I don't want to remove all of those entries. Let's take a look at the distribution of ages, then impute with the average or median ages.

```{r}
ggplot(data=train, aes(x=Age)) +
geom_histogram(color="darkblue", fill="lightblue", bins = 50, na.rm =TRUE) +
 geom_vline(mapping = aes(xintercept=mean(Age, na.rm=TRUE)), color='red', linetype='dashed', size=1) +
 geom_vline(mapping = aes(xintercept=median(Age, na.rm=TRUE)), color='green', linetype='dashed', size=1)
```
Since this is a Poison distribution we'll use the median. The median is also visually a little more in the middle.

Let's also check to see if some of the features are correlated.
We'll look first at Fare and Class
```{r}
ggplot(data=train, mapping = aes(x=Pclass, y=Fare))+
  geom_boxplot() +
  geom_point(aes(y=Fare, x=Pclass, color = Survived), alpha = 0.5) 
```
















