---
title: "Titanic Dataset - Exploration with R"
output: html_notebook
---

Get the required packages:
```{r}
require(ggplot2)
```

Load the data into R:
```{r}
train <- read.csv("train.csv", stringsAsFactors = F) #For cleaning purposes let's not convert strings to factors.
test <- read.csv("test.csv", stringsAsFactors = F)
```

Let's take a look at what is in the train and test data sets.
```{r}
str(train)
```

Let's see how many missing values there are in each of the columns.
```{r}
for (n in names(train)){
  cat(sprintf("%s missing %i values.\n", n, sum(is.na(train[n]))))
}
```
It appears that the "Age" column has a significant number of missing values.
Luckily however there are not missing values in the other columns.

Before we remove or impute the missing values, let's just make sure there is not a correlation between the missing values and the predictions.
```{r}
#Let's get a baseline for what ratio survived.
sum(train$Survived)/length(train$Survived)
#Now do the same for the "NA" Ages.
train_na <- subset(train, is.na(train$Age))
sum(train_na$Survived)/length(train_na$Survived)
```






